---
title: "Airbnb Pricing Tool"
author: "Team R"
date: "31 Januar 2018"
output:
  html_document:
    css: ../AnalyticsStyles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: ../AnalyticsStyles/default.sty
always_allow_html: yes
---

<!-- rmarkdown::render("AirbnbProject/Airbnb_Pricing_TeamR_MASTER.rmd") -->

```{r echo=FALSE, message=FALSE}
make_pdf_file = 0 # SET THIS TO 1 IF WE COMPILE PDF FILE, 0 OTHERWISE (FOR HTML)

source("../AnalyticsLibraries/library.R")
source("../AnalyticsLibraries/heatmapOutput.R")

# Package options
ggthemr('fresh')  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.2)
options(knitr.kable.NA = '')
```

```{r echo=FALSE, message=FALSE}
# Please ENTER the filename
datafile_name = "Data/tomslee_airbnb_amsterdam_1476_2017-07-22.csv"
PricingData <- read.csv(datafile_name)
# We turn the data into data.matrix class so that we can easier manipulate it
PricingData <- data.matrix(PricingData)

# Please ENTER the dependent variable.
dependent_variable = 14 # i.e Price

# Please ENTER the attributes to use as independent variables. 
independent_variables = c(1:13,18:19) # use all the available attributes

max_data_report = 10
```

# The Business Context
Airbnb was founded in 2008 by *Brian Chesky*, *Joe Gebbia*, and *Nathan Blecharczyk* as *AirBed & Breakfast*, an online marketplace and hospitality service for short-term lodging. Over the past years, the share of professional hospitality providers has significantly increased and is now *crowding out* the private providers, threatening Airbnb's value proposition of *unique design and personal touch*. Within this context, the marketing department wants to run a campaign to attract more private providers. To do this, they requested the analytics department to create a tool that helps interested landlords to understand how much money they could earn with their respective apartment. 

As a pilot, Amsterdam was chosen because of the "Airbnb friendly" policy of the local regulators and the high number of short-term visitors. The proposed solution, however, is designed to be city-independent and therefore, an easily replicable process was designed, using *.rmd-Files* and *Github*.


<hr>\clearpage

# The Data
(Data source: http://tomslee.net/airbnb-data-collection-get-the-data. We acknowledge the following: All material is copyright Tom Slee, licensed under a Creative Commons Attribution-NonCommercial 2.5 Canada License.)

The data is collected from the official Airbnb website

Name                       | Description
:--------------------------|:--------------------------------------------------------------------
ID                         | ID of each client
LIMIT_BAL                  | Amount of given credit in NT dollars (includes individual and family/supplementary credit)
SEX                        | Gender (1=male, 2=female)
EDUCATION                  | (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
MARRIAGE                   | Marital status (1=married, 2=single, 3=others)
AGE                        | Age in years
PAY_0                      | Repayment status in September, 2005 (-2=no consumption, -1=pay duly, 0=the use of revolving credit, 1=payment delay for one month, 2=payment delay for two months, ... 8=payment delay for eight months, 9=payment delay for nine months and above)
PAY_2                      | Repayment status in August, 2005 (scale same as above)
PAY_3                      | Repayment status in July, 2005 (scale same as above)
PAY_4                      | Repayment status in June, 2005 (scale same as above)
PAY_5                      | Repayment status in May, 2005 (scale same as above)
PAY_6                      | Repayment status in April, 2005 (scale same as above)
BILL_AMT1                  | Amount of bill statement in September, 2005 (NT dollar)
BILL_AMT2                  | Amount of bill statement in August, 2005 (NT dollar)
BILL_AMT3                  | Amount of bill statement in July, 2005 (NT dollar)
BILL_AMT4                  | Amount of bill statement in June, 2005 (NT dollar)
BILL_AMT5                  | Amount of bill statement in May, 2005 (NT dollar)
BILL_AMT6                  | Amount of bill statement in April, 2005 (NT dollar)
PAY_AMT1                   | Amount of previous payment in September, 2005 (NT dollar)
PAY_AMT2                   | Amount of previous payment in August, 2005 (NT dollar)
PAY_AMT3                   | Amount of previous payment in July, 2005 (NT dollar)
PAY_AMT4                   | Amount of previous payment in June, 2005 (NT dollar)
PAY_AMT5                   | Amount of previous payment in May, 2005 (NT dollar)
PAY_AMT6                   | Amount of previous payment in April, 2005 (NT dollar)
default.payment.next.month | Default payment (1=yes, 0=no)

Let's look into the data for a few customers. This is how the first `r min(max_data_report, nrow(PricingData))` out of the total of `r nrow(PricingData)` rows look like (transposed, for convenience):

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
knitr::kable({
  df <- t(head(round(PricingData[,independent_variables],2), max_data_report))
  colnames(df) <- sprintf("%02d", 1:ncol(df))
  df
})
```
